---
title: "sncrna_june"
author: "Chishan Burch"
date: "2025-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("readxl")
#install.packages("openxlsx")
#install.packages("ggplot2")

#install.packages("DESeq2")
#install.packages("dplyr")
#install.packages("tibble")
#install.packages("ggrepel")
#install.packages("edgeR")

library(openxlsx)
library(dplyr)
library(tibble)
library(DESeq2)
library(ggplot2)
library(openxlsx)
library(readxl)
library(ggrepel)
library(edgeR)

```

```{r}
#A2 <- read.xlsx("./1_data/sncRNA_june/A2_june.xlsx")
#A3 <- read.xlsx("./1_data/sncRNA_june/A3_june.xlsx")
#A4 <- read.xlsx("./1_data/sncRNA_june/A4_june.xlsx")
#A5 <- read.xlsx("./1_data/sncRNA_june/A5_june.xlsx")

#P1 <- read.xlsx("./1_data/sncRNA_june/P1_june.xlsx")
#P3 <- read.xlsx("./1_data/sncRNA_june/P3_june.xlsx")
#P4 <- read.xlsx("./1_data/sncRNA_june/P4_june.xlsx")
#P5 <- read.xlsx("./1_data/sncRNA_june/P5_june.xlsx")

# The data is not in the structure I am used to so I need to re-format it before I can do my pipeline 
#### Here I am creating an example of the final structure I want my data to be in so I can prompt chatGPT to give me the correct output

#sRNA.id <- c("example_pir1","example_pir2", "example_pir3", "example_pir4", "example_pir5", "example_pir6") 
#Type <- c("example_type", "example_type","example_type", "example_type", "example_type", "example_type")
#Description <- c("example_desc", "example_desc","example_desc", "example_desc","example_desc", "example_desc")
#A2 <- c(1, 66, 35, 45, 2, 34) 
#A3 <- c(8, 234, 35, 283, 289, 34) 
#A4 <- c(1, 200, 445, 283, 289, 34) 
#A5 <- c(445, 0, 35, 3, 289, 34) 
#P1 <- c(1, 200, 35, 0, 43, 34) 
#P3 <- c(1, 2, 35, 283, 289, 34) 
#P4 <- c(56, 33, 35, 283, 289, 34) 
#P5 <- c(1, 200, 35, 283, 9, 34) 
            
#example_df <- data.frame(sRNA.id, Type, Description, A2, A3, A4, A5, P1, P3, P4, P5, stringsAsFactors = FALSE)

# Asked ChatGPT "My data frames are named A2, A3, A4, A5, P1, P3, P4, P5. How can I combine them to look like my data frame example_df?"
###
```


```{r}
# I was having trouble with duplicated rows, so we will try combining the sRNA.id and Type columns

#library(tidyr)

# Combine the 'sRNA.id' and 'Type' columns with 'Type' after 'sRNA.id'
#A2 <- A2 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#A3 <- A3 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#A4 <- A4 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#A5 <- A5 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#P1 <- P1 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#P3 <- P3 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#P4 <- P4 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)

#P5 <- P5 %>%
#  unite("sRNA_Type", sRNA.id, Type, sep = " ", remove = FALSE)
```

```{r}
# Step 1: Store all data frames in a list
#samples <- list(A2 = A2, A3 = A3, A4 = A4, A5 = A5, P1 = P1, P3 = P3, P4 = P4, P5 = P5)

# Step 2: Keep only `sRNA_Type` and sample-specific column, process into a new list
#processed_samples <- lapply(names(samples), function(sample_name) {
#  df <- samples[[sample_name]]  # Access the original data frame
  
  # Check if the sample column exists
#  if (!sample_name %in% colnames(df)) {
#    stop(paste("Missing column:", sample_name, "in data frame."))
#  }
  
  # Select and rename columns
#  df <- df[, c("sRNA_Type", sample_name)]
#  colnames(df)[2] <- "sample_value"  # Standardize column name for merging
#  return(df)
#})

# Step 3: Merge all data frames by `sRNA_Type`
#merged_data <- processed_samples[[1]]  # Start with the first data frame

#for (i in 2:length(processed_samples)) {
  # Retrieve the current sample data frame
#  current_sample <- processed_samples[[i]]
#  sample_name <- names(samples)[i]
  
  # Rename `sample_value` dynamically
#  colnames(current_sample)[2] <- sample_name  # Rename to the sample's name
  
  # Merge with the existing merged data
#  merged_data <- merge(merged_data, current_sample, by = "sRNA_Type", all = TRUE)
#}

# Step 4: Add `Type` and `Description` columns from A2
#merged_data <- merge(
#  merged_data,
#  A2[, c("sRNA_Type", "sRNA.id", "Type", "Description")],
#  by = "sRNA_Type",
#  all.x = TRUE
#)

# Step 5: Reorder columns
#final_columns <- c("sRNA_Type", "sRNA.id", "Type", "Description", names(samples))
#merged_data <- merged_data[, final_columns]

# Step 6: Replace NA with 0 (optional)
#merged_data[is.na(merged_data)] <- 0

# Step 7: View the final merged data
#print(merged_data)

# Renaming column 2
#merged_data <- merged_data %>%
#  rename("sample_value" = "A2")

#write.csv(merged_data, file ="./1_data/sncRNA_june/raw_UMI_counts_sncRNA_june.csv")
```

QC and filter raw data
```{r include=FALSE}
df <- read.csv(file = "./1_data/sncRNA_june/raw_UMI_counts_sncRNA_june.csv")

min.counts <- 5
min.samples <- 3

# Removing uneeded columns
df <- df[, -c(1, 11, 12, 13)]

# Make the sncRNA ids into rownames
df <- column_to_rownames(df, var = "sRNA_Type")

 df <- df %>%
   filter(rowSums(. >= min.counts) >= min.samples)

    if (any(is.na(df))) {
      df <- df[complete.cases(df), ]
    }
```

Top 6 rows (omitted 1603 rows)
```{r}
head(df)

```


# Principal component analysis (PCA) on filtered small RNA counts
```{r}

# A = Acrylamide, P = Control
sample_metadata <- read.csv("./1_data/sncRNA_sample_metadata.csv")

sample_metadata$Treatment_group <- as.factor(sample_metadata$Treatment_group)
levels(sample_metadata$Treatment_group)

level_colors <- c("Acr" = "#994455", "Control" = "#0077BB")



PCA <- prcomp(t(df), center = TRUE)
#######
# Count the number of duplicate values in the column 'xcolumn'
#num_duplicates <- sum(duplicated(df$sRNA_Type))

# Print the result
#print(num_duplicates)



####
pca_data <- data.frame(PCA$x)
pca_data$SampleID <- rownames(pca_data)



# Ensure SampleID column exists in sample_metadata and matches pca_data
sample_metadata$SampleID <- as.character(sample_metadata$SampleID)

# Merge Treatment_group into pca_data using SampleID as the key
pca_data <- merge(pca_data, sample_metadata[, c("SampleID", "Treatment_group")], by = "SampleID", all.x = TRUE)



# Vareince explained
variance_explained <- 100 * PCA$sdev^2 / sum(PCA$sdev^2)


# Plotting
PCA_sncRNAs <- ggplot(pca_data, aes(x = PC1, y = PC2, color = Treatment_group, label = SampleID)) +
  geom_point(size = 4, shape = 21, fill = "white", stroke = 1) +
  geom_label_repel(size = 3, box.padding = 0.25, label.padding = 0.5) +
  scale_color_manual(values = level_colors) +
  labs(
    title = "Dimensionality Reduction Highlights Transcriptomic Shifts in small RNAs",
    color = "Treatment group", 
    x = paste0("PC1 (", round(variance_explained[1], 1), "%)"), 
    y = paste0("PC2 (", round(variance_explained[2], 1), "%)")
  ) +
  guides(color = guide_legend(title = "Treatment group")) +
  theme_minimal()

plot(PCA_sncRNAs)
# Save the plot as a PNG
#ggsave(
#  filename = "./3_results/PCA1_sncRNAs.png", 
#  plot = PCA_sncRNAs,           
#  width = 8,                    
#  height = 6,                   
#  dpi = 300                     
#)

```


# DE analysis and volcano plots
```{r, include=FALSE}
# Relevel the Treatment_group factor
sample_metadata$Treatment_group <- relevel(sample_metadata$Treatment_group, ref = "Control")

# DE analysis
dds <- DESeqDataSetFromMatrix(countData = df,
                              colData = sample_metadata,
                              design = ~Treatment_group)
dds <- DESeq(dds)

# Results and filtering
res <- results(dds, contrast = c("Treatment_group", "Acr", "Control")) 
res_df <- as.data.frame(res)
res_df <- na.omit(res_df)  # Remove rows with NA values


# Variance Stabilizing Transformation and PCA
vsd <- varianceStabilizingTransformation(dds, blind = TRUE)
vst_matrix <- assay(vsd)

pca <- prcomp(t(vst_matrix))
plot(pca$x[, 1], pca$x[, 2], main = "PCA of VST-transformed small RNA data", 
     xlab = "PC1", ylab = "PC2")

# Annotate significant DE results
res_df <- res_df %>%
  mutate(sign_DE = if_else(padj < 0.05 & log2FoldChange < -0.585, "Sign_Down",
                           if_else(padj < 0.05 & log2FoldChange > 0.585, "Sign_Up", "NS")))

# Add a -log10 column for volcano plot
res_df <- res_df %>% mutate(log_padj = -log10(padj))

# Save results to CSV
#write.csv(res_df, file = "./1_data/sncRNA_june/deseq_sncrnas_june.csv")

# Volcano Plot
colours_1 <- c("Sign_Up" = "#994455", "Sign_Down" = "#0077BB", "NS" = "#BBBBBB") 
opacity <- c("Sign_Up" = 1, "Sign_Down" = 1, "NS" = 0.5)

plot_1 <- res_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))


```


```{r, include=FALSE}
# Use grep() to separate the different sncRNA types **plot if asked

# Mature miRNA
rows <- grep("mature", rownames(res_df))
mature_miRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(mature_miRNA_df)

# Volcano Plot

plot_2 <- mature_miRNA_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))

# Display plot
plot(plot_2)

# Hairpin miRNA
rows <- grep("hairpin", rownames(res_df))
hairpin_miRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(hairpin_miRNA_df)

plot_3 <- hairpin_miRNA_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))

# Display plot
plot(plot_3)

# piRNA 
rows <- grep("piRNA", rownames(res_df))
piRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(piRNA_df)


plot_4 <- piRNA_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))

# Display plot
plot(plot_4)


# rRNA
rows <- grep("rRNA", rownames(res_df))
rRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(rRNA_df)

plot_5 <- rRNA_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))

# Display plot
plot(plot_5)


# snoRNA
rows <- grep("snoRNA", rownames(res_df))
snoRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(snoRNA_df)

plot_6 <- snoRNA_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))

plot(plot_6)


# snRNA
rows <- grep("snRNA", rownames(res_df))
snRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(snRNA_df)


rows <- grep("snoRNA", rownames(res_df))
snoRNA_df <- res_df[rows, , drop = FALSE] # Ensure drop=FALSE to keep full rows
print(snoRNA_df)

plot_6 <- snRNA_df %>%
  filter(!is.na(sign_DE)) %>%
  ggplot(aes(x = log2FoldChange, y = log_padj, col = sign_DE, alpha = sign_DE, fill = sign_DE)) +
  geom_point(size = 5, shape = 21, color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", size = 0.4, col = "#3A3B3C") +
  scale_color_manual(values = colours_1) + 
  scale_fill_manual(values = colours_1) +
  scale_alpha_manual(values = opacity) + 
  theme_bw() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
        legend.text = element_text(size = 14), legend.title = element_text(size = 14))

plot(plot_6)



```


```{r}
# mature miRNA
plot(plot_1)
```
Take the validated targets of downregulated miRNAs from multimiR, and see of those same genes are upregulated in the TEtranscripts dataset.Use volcano plot. 
Take sig down miRNA tagrtes
Plot them on volcano plot, plot all miRNA targets as blue. Every gene will be grey exceot the targets of downregulated miRNAs
Do the same on a separate plot with upregulated miRNA targets.



```{r}
# hairpin miRNA
plot(plot_2)
```

```{r}
# piRNA
plot(plot_3)
```

```{r}
# rRNA
plot(plot_4)
```

```{r}
# snoRNA
plot(plot_5)
```

```{r}
# snRNA
plot(plot_6)
```

mRNA target prediction
```{r}

# We want to identify mRNAs which are targeted by upregulated and downregulated miRNAs

# Load libraries
#library(openxlsx)
library(data.table)
library(dplyr)
#library(tibble)

library(ggplot2)
library(ggvenn)

#BiocManager::install("multiMiR")


# Warning in install.packages :
 # package ‘multiMiR’ is not available for this version of R

#A version of this package for your version of R might be available elsewhere,
#see the ideas at
#https://cran.r-project.org/doc/manuals/r-patched/R-admin.html#Installing-packages

library(multiMiR)
#library(edgeR)
#library(limma)

# Remove 'mature' from row names
rownames(mature_miRNA_df) <- gsub(" mature", "", rownames(mature_miRNA_df)) #mature
rownames(hairpin_miRNA_df) <- gsub(" hairpin", "", rownames(hairpin_miRNA_df)) #hairpin

mature_miRNA_df <- rownames_to_column(mature_miRNA_df, "miRNA_ids")
hairpin_miRNA_df <- rownames_to_column(hairpin_miRNA_df, "miRNA_ids")

# Follow 5_miR_target_prediction.R from placement. (Up to line 134)
significant_miRNAs <- mature_miRNA_df %>%
  filter(padj < 0.05)
significant_miRNAs_hairpins <- hairpin_miRNA_df %>%
  filter(padj < 0.05)

# Create a df of all the mature miRNAs with padj < 0.05, and a df of all the mature miRNAs to pass the general filters.

# significant miRNAs are here
print(significant_miRNAs)
# all the miRNAs are here
print(mature_miRNA_df)


# Upregulated and downregulated miRNAs
miRNA_UP <- significant_miRNAs %>%
filter(log2FoldChange > 0)

miRNA_DOWN <- significant_miRNAs %>%
  filter(log2FoldChange < 0)

#hairpins_UP <- significant_miRNAs_hairpins %>%
#  filter(log2FoldChange > 0)

#hairpins_DOWN <- significant_miRNAs_hairpins %>%
#  filter(log2FoldChange < 0)


# Get vectors of miRNA ids and remove empty strings

mature_up <- miRNA_UP$miRNA_ids
mature_up <- mature_up[mature_up != ""] # REMOVING EMPTY STRINGS IS IMPORTANT

mature_down <- miRNA_DOWN$miRNA_ids
mature_down <- mature_down[mature_down != ""]

#hairpins_up <- hairpins_UP$miRNA_ids
#hairpins_up <- hairpins_up[hairpins_up != ""]

#hairpins_down <- hairpins_DOWN$miRNA_ids
#hairpins_down <- hairpins_down[hairpins_down != ""]

# need to specify org = "mmu" every time the function is used


# MATURE MIRNA UP PREDICTED AND VALIDATED ################
# MUP = mature, upregulated, predicted
# MUV = mature, downregulated, validated
#multimiR_MUP <- get_multimir(org = "mmu",
#                                 mirna = mature_up,
#                                 table = c("predicted"),
#                                 summary = TRUE)
multimiR_MUV <- get_multimir(org = "mmu",
                                 mirna = mature_up,
                                 table = c("validated"),
                                 summary = TRUE)
# MATURE MIRNA DOWN PREDICTED AND VALIDATED ################
#multimiR_MDP <- get_multimir(org = "mmu",
#                                 mirna = mature_down,
#                                 table = c("predicted"),
#                                 summary = TRUE)
multimiR_MDV <- get_multimir(org = "mmu",
                                 mirna = mature_down,
                                 table = c("validated"),
                                 summary = TRUE)
# HAIRPIN MIRNA UP PREDICTED AND VALIDATED ################
#multimiR_HUP <- get_multimir(org = "mmu",
#                                 mirna = hairpins_up,
#                                 table = c("predicted"),
#                                 summary = TRUE)
#multimiR_HUV <- get_multimir(org = "mmu",
#                                 mirna = hairpins_up,
#                                 table = c("validated"),
#                                 summary = TRUE)
# HAIRPIN MIRNA DOWN PREDICTED AND VALIDATED ################
#multimiR_HDP <- get_multimir(org = "mmu",
#                                 mirna = hairpins_down,
#                                 table = c("predicted"),
#                                 summary = TRUE)
#multimiR_HDV <- get_multimir(org = "mmu",
#                                 mirna = hairpins_down,
#                                 table = c("validated"),
#                                 summary = TRUE)
## omgosh that was alot

# Now pass each one as a data frame and save it
#mature upregulated
#multimiR_MUP <- as.data.frame(multimiR_MUP@data)
multimiR_MUV <- as.data.frame(multimiR_MUV@data)

# mature downregulated
#multimiR_MDP <- as.data.frame(multimiR_MDP@data)
multimiR_MDV <- as.data.frame(multimiR_MDV@data)

# hairpin upregulated
#multimiR_HUP <- as.data.frame(multimiR_HUP@data)
#multimiR_HUV <- as.data.frame(multimiR_HUV@data)

# hairpin downregulated
#multimiR_HDP <- as.data.frame(multimiR_HDP@data)
#multimiR_HDV <- as.data.frame(multimiR_HDV@data)

# Save them, girlqueen
#write.csv(multimiR_MUP, file = "./3_results/multimiR_mature_mirnas_up_predicted.csv")
#write.csv(multimiR_MUV, file = "./3_results/multimiR_mature_mirnas_up_validated.csv")

#write.csv(multimiR_MDP, file = "./3_results/multimiR_mature_mirnas_down_predicted.csv")
#write.csv(multimiR_MDV, file = "./3_results/multimiR_mature_mirnas_down_validated.csv")

#write.csv(multimiR_HUP, file = "./3_results/multimiR_hairpin_mirnas_up_predicted.csv")
#write.csv(multimiR_HUV, file = "./3_results/multimiR_hairpin_mirnas_up_validated.csv")

#write.csv(multimiR_HDP, file = "./3_results/multimiR_hairpin_mirnas_down_predicted.csv")
#write.csv(multimiR_HDV, file = "./3_results/multimiR_hairpin_mirnas_down_validated.csv")

# whip and nae nae


# Take identified miRNA targets and identify them in the mRNA volcano plot/dataset. Only use the validated targets for this.
# On the volcano plot with all genes, highlight the (validated) genes targeted by miRNAs. Have different colours for up and down regulated targets. 


```



KEGG analysis of mRNA targets
```{r}

# Combine up and down into a single vector before doing GO-ORA and KEGG because all miRNAs can be up or downregulated at different times. Doing it like 
# this strengthnes the results

# Just do the mature ones for the scope of honours thesis

library(org.Mm.eg.db)
library(AnnotationDbi)

# For this I follow TEtranscripts_april.Rmd


# Obtain vectors of entrez ids 

# I'm gonna overwrite the multimir results to preserve memory since they are already saved as csv files by this point.  

# matrue upregulated
multimiR_MUV <- multimiR_MUV

# mature downregulated
multimiR_MDV <- multimiR_MDV

head(multimiR_MDV)
head(multimiR_MUV)

# Combine both data frames into one
combined_df <- rbind(multimiR_MUV, multimiR_MDV)

# Filter out duplicate  entries (keeping the first occurrence)
unique_df <- combined_df[!duplicated(combined_df$target_entrez), ]

# If you just want the vector of unique mature_mirna_id values:
unique_multimiR_ids <- unique(combined_df$target_entrez)


# HIT LIST
head(unique_multimiR_ids)

# FOR the background list, use the targets of miRNAs meeting basic QC (pre DESeq), because the multimiR results are made from miRNAs 
# which stand out post-DESEq2
###############


genes <- fread(file = "./1_data/Transcriptomics/TE_and_gene_counts_Updated_and_Raw.txt")

# Isolating the gene names instead of the ensembl ids.
background_list <- genes$ensembl.TE

# BACKGROUND LIST 
head(background_list)

# Gotta convert background list to ensembl
#BiocManager::install("org.Mm.eg.db")
library(org.Mm.eg.db)  
organism <- org.Mm.eg.db    

library(clusterProfiler)

# Convert gene names to KEGG IDs
    KEGG_background_list <- bitr(
      background_list,
      fromType = "ENSEMBL",  # Source ID type (e.g., HGNC symbol)
      toType = "ENTREZID",      # Target ID type
      OrgDb = organism  
    )

KEGG_background_list <- KEGG_background_list$ENTREZID

# Check if there is some overlap between the hit list and background list
length(intersect(unique_multimiR_ids, KEGG_background_list)) 

KEGG_enrichment_results <- enrichKEGG(
  gene = unique_multimiR_ids,
  universe = KEGG_background_list,
  organism = "mmu",  # Mouse
  keyType = "kegg",  # Key type in KEGG
  pvalueCutoff = 0.05,
  qvalueCutoff = 0.05
)

print(KEGG_enrichment_results)



#KEGG_barplot <- barplot(KEGG_enrichment_results, showCategory = 20)
#KEGG_dotplot <- dotplot(KEGG_enrichment_results, color = "pvalue")

# Save KEGG barplot
#png("KEGG_barplot.png", width = 1200, height = 900, res = 150)
#barplot(KEGG_enrichment_results, showCategory = 20)
#dev.off()

# Save KEGG dotplot
#png("KEGG_dotplot.png", width = 1200, height = 900, res = 150)
#dotplot(KEGG_enrichment_results, color = "pvalue")
#dev.off()



#KEGG_results <- as.data.frame(KEGG_enrichment_results@result)

#write.csv(KEGG_results, file = "./3_results/sncRNA_July_KEGG.csv")
############################################################################## BACKGROUND LIST IS THE TETRANSCRPTS ENSUMBL IDS TURNED INTO ENTREZ
## OMGGG I NEEDED THE ENTREZ IDS ARRGHH#################




```


GO analysis of mRNA targets.
```{r}

# Produced during KEGG analysis section
# ENSEMBL HIT LIST
unique_multimiR_ids_ensembl <- combined_df$target_ensembl
# ENSEMB BACKGROUND LIST
head(background_list)

go_enrich <- enrichGO(gene = unique_multimiR_ids_ensembl,
                     universe = background_list,
                      OrgDb = organism, 
                      keyType = 'ENSEMBL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)


# Extract and save results table

go_results <- go_enrich@result
go_results$GeneRatio <- gsub("/","//",go_results$GeneRatio)
go_results$BgRatio <- gsub("/","//",go_results$BgRatio)


#write.xlsx(go_results, file = "./3_results/sncRNA_GOenrich.xlsx")

# Save table

# Step 4 - Plots
## Bar plot
barplot <- barplot(go_enrich,
        title = "GO Biological Pathways")

plot(barplot)
#ggsave(file = "./2_figures/sncRNA_GO_barplot.png")

## Dot plot
enrichplot::dotplot(go_enrich)

#ggsave(file = "./2_figures/sncRNA_GO_dotplot.png")

## To remove/account for redundant parent pathways (all those terms belonging to the same general functional group)
edox2 <- pairwise_termsim(go_enrich)

### Tree plots
p1 <- treeplot(edox2, offset = 10)
plot(p1)

#ggsave("./2_figures/sncRNA_GO_treeplot1.png", plot = p1, width = 15, height = 6, units = "in", dpi = 300)

p2 <- treeplot(edox2, hclust_method = "average", offset = 10)
plot(p2)

#ggsave("./2_figures/sncRNA_GO_treeplot2.png", plot = p2, width = 15, height = 6, units = "in", dpi = 300)

p3 <- treeplot(edox2, offset = 7,
               nCluster = 5, # default is 5. Changes number of groups clustered
               showCategory = 10)  # Number of 'top terms' that will be grouped and displayed in the plot
plot(p3)

#ggsave("./2_figures/sncRNA_GO_treeplot3.png", plot = p3, width = 15, height = 6, units = "in", dpi = 300)


```






Perform TMM normalisation of sncRNA

REDO THE TMM for sncRNA and for mRNA
```{r}
counts <- df

# # Assuming 'counts' is a matrix of raw count data with genes as rows and samples as columns
# # Create a DGEList object
y <- DGEList(counts = counts)

# # Perform TMM normalization
y <- calcNormFactors(y)

# # Get the normalised counts
TMM_matrix <- cpm(y, prior.count = 1, normalized.lib.sizes = TRUE)

# # Save them. They will be used for PCA plots.
#write.csv(TMM_matrix, file = "./3_results/sncRNA_june_TMM_matrix.csv", row.names = TRUE)


```



PCA plot of TMM
```{r}

# 3. PCA plots – 1 x Gene, 1 x TE, 1 x ALL 

# Start plotting
# Perform PCA
sncRNA_PCA <- prcomp(t(TMM_matrix), center = FALSE)

# Create PCA data frame
pca_data <- data.frame(sncRNA_PCA$x)
pca_data$SampleID <- rownames(pca_data)

# Ensure SampleID is a common column for merging
if ("SampleID" %in% colnames(pca_data) & "SampleID" %in% colnames(sample_metadata)) {
  pca_data <- merge(pca_data, sample_metadata, by = "SampleID")
} else {
  stop("Column 'SampleID' not found in one of the data frames.")
}

# Set colors for the groups
#level_colors <- c("Acr" = "#994455", "Control" = "#0077BB")


# Plot 
sncRNA_PCA <- prcomp(pca_data[, c("PC1", "PC2")], center = TRUE, scale. = TRUE)

# Extract percentage of variance explained
variance_explained <- 100 * sncRNA_PCA$sdev^2 / sum(sncRNA_PCA$sdev^2)


# Create PCA plot using ggplot2
sncRNA_PCA <- ggplot(pca_data, aes(x = PC1, y = PC2, color = Treatment_group, label = SampleID)) +
  geom_point(size = 4, shape = 21, fill = "white", stroke = 1) +  # Add shading with `fill` and `stroke` for border thickness
  geom_label_repel(size = 3, box.padding = 0.25, label.padding = 0.5) +
  scale_color_manual(values = level_colors) +
  labs(title = "Dimensionality Reduction Highlights Transcriptomic Shifts in SV (all sncRNAs)",
       color = "Treatment group", 
       x = paste0("PC1 (", round(variance_explained[1], 1), "%)"), 
       y = paste0("PC2 (", round(variance_explained[2], 1), "%)")) +
  guides(color = guide_legend(title = "Treatment group")) +
  theme_minimal() 

plot(sncRNA_PCA)
```


# Boxplot of TMM normalised small RNA counts (log2)
```{r, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)

# Convert matrix to data frame and reshape it
TMM_df <- as.data.frame(TMM_matrix) %>%
  tibble::rownames_to_column(var = "Gene") %>%  # or "id", depending on preference
  pivot_longer(
    cols = -Gene,  # pivot all columns except Gene name
    names_to = "Sample",
    values_to = "Expression"
  )

# Plot for TMM sncRNAs
TMM_boxplot <- ggplot(TMM_df, aes(x = Sample, y = log2(Expression + 1))) +
  geom_boxplot() +
  labs(title = "Boxplot for TMM small RNAs", x = "Sample", y = "Expression") +
  theme_minimal()

```

*** Do a boxplot for raw counts as well for comparison
```{r}
# NO IDEA WHY THIS ISNT WORKING ARGHH

# Convert matrix to data frame and reshape it
raw_df <- as.data.frame(df) %>%
  tibble::rownames_to_column(var = "Gene") %>%  # or "id", depending on preference
  pivot_longer(
    cols = -Gene,  # pivot all columns except Gene name
    names_to = "Sample",
    values_to = "Expression"
  )

# Plot for raw sncRNA counts
boxplot <- ggplot(raw_df, aes(x = Sample, y = log2(Expression + 1))) +
  geom_boxplot() +
  labs(title = "Boxplot for small RNAs (raw counts)", x = "Sample", y = "Expression") +
  theme_minimal()

```

# Boxplot

```{r}
plot(boxplot)
```
# TMM boxplot

It's only slightly different
```{r}
plot(TMM_boxplot)
```
