---
title: "sncRNA analysis of acrylamide-treated vs control seminal vesicles (SPORTS package)"
author: "Chishan Burch"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(openxlsx)
```

```{r, include=FALSE}
#Threshold for FDR which is the same as p-adjusted
FDR_threshold <- 0.05
#Declare graph colours and factor values
palette <- c("Upreg" = "#D55E00", "Downreg" = "#0072B2", "NS" = "#808080")
```

Horizonal dashed line represents FDR = -log10(0.05)

Vertical dashed lines represent log2FC = -0.5833 and 0.5833

**miRNA dysregulation**

```{r, include=FALSE}
#Start with miRNA
miRNA <- fread("./1_data/miRNA_SCH_glmLRT_FC.csv")

#Adding column for significance and up/down regulation (NS means not significant)
miRNA <- miRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0.5833, "Upreg", "NS")
                          
        )) 
#Adding column for -log10FDR
miRNA <- miRNA %>% mutate(log_FDR = -log10(FDR))

#Adding column for up and down reg
miRNA$Significance <- factor(miRNA$Significance, levels = c("Upreg","Downreg", "NS"))
 
#plot miRNAs
miRNA_plot <- ggplot(data = miRNA, aes(x =logFC, y= log_FDR)) + 
  geom_point(size = 2.5) + theme_minimal() +
  geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(label = Annotation, colour = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("miRNA") 

```

```{r}
miRNA_plot
```

```{r, include=FALSE}
DE_miRNAs <- miRNA %>% select(Annotation, logFC, log_FDR)
DE_miRNAs_desc <- DE_miRNAs %>% arrange(desc(logFC))
DE_miRNAs_asc <- DE_miRNAs %>% arrange((logFC))

miRNA_FDRs <- DE_miRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated miRNAs (highest to lowest FC):

```{r}
head(DE_miRNAs_desc)
```

Here are the most downregulated miRNAs (lowest to highest FC):

```{r}
head(DE_miRNAs_asc)
```

Here is a list of miRNAs organised from highest to lowest FDR:

```{r}
head(miRNA_FDRs)
```

**tRNA dysregulation**

```{r, include=FALSE}
tRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "tRNA")

#Adding column for significance and up/down regulation (NS means not significant)
tRNA <- tRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 
#Adding column for -log10FDR
tRNA <- tRNA %>% mutate(log_FDR = -log10(FDR))

tRNA$Significance <- factor(tRNA$Significance, levels = c("Upreg","Downreg", "NS"))

#plot tRNAs
tRNA_plot <- ggplot(data = tRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal() +
  geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("tRNA")

```

```{r}
tRNA_plot
```

```{r, include=FALSE}
DE_tRNAs <- tRNA %>% select(Annotation, logFC, log_FDR)
DE_tRNAs_desc <- DE_tRNAs %>% arrange(desc(logFC))
DE_tRNAs_asc <- DE_tRNAs %>% arrange((logFC))

tRNA_FDRs <- DE_tRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated tRNAs (highest to lowest FC):

```{r}
head(DE_tRNAs_desc)
```

Here are the most downregulated tRNAs (lowest to highest FC):

```{r}
head(DE_tRNAs_asc)
```

Here is a list of tRNAs organised from highest to lowest FDR:

```{r}
head(tRNA_FDRs)
```

**rRNA dysregulation**

```{r, include=FALSE}
#rRNA
rRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "rRNA")

#Adding column for significance and up/down regulation (NS means not significant)
rRNA <- rRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 
#Adding column for -log10FDR
rRNA <- rRNA %>% mutate(log_FDR = -log10(FDR))

rRNA$Significance <- factor(rRNA$Significance, levels = c("Upreg","Downreg", "NS"))

#plot rRNAs
rRNA_plot <- ggplot(data = rRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("rRNA")
```

```{r}
rRNA_plot
```

```{r, include=FALSE}
DE_rRNAs <- rRNA %>% select(Annotation, logFC, log_FDR)
DE_rRNAs_desc <- DE_rRNAs %>% arrange(desc(logFC))
DE_rRNAs_asc <- DE_rRNAs %>% arrange((logFC))

rRNA_FDRs <- DE_rRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated rRNAs (highest to lowest FC):

```{r}
head(DE_rRNAs_desc)
```

Here are the most downregulated rRNAs (lowest to highest FC):

```{r}
head(DE_rRNAs_asc)
```

Here is a list of rRNAs organised from highest to lowest FDR:

```{r}
head(rRNA_FDRs)
```

**lincRNA dysregulation**

```{r, include=FALSE}
lincRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "lincRNA")

lincRNA <- lincRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 
lincRNA <- lincRNA %>% mutate(log_FDR = -log10(FDR))

lincRNA$Significance <- factor(lincRNA$Significance, levels = c("Upreg","Downreg", "NS"))


lincRNA_plot <- ggplot(data = lincRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("lincRNA")
```

```{r}
lincRNA_plot
```

```{r, include=FALSE}
DE_lincRNAs <- lincRNA %>% select(Annotation, logFC, log_FDR)
DE_lincRNAs_desc <- DE_lincRNAs %>% arrange(desc(logFC))
DE_lincRNAs_asc <- DE_lincRNAs %>% arrange((logFC))

lincRNA_FDRs <- DE_lincRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated lincRNAs (highest to lowest FC):

```{r}
head(DE_lincRNAs_desc)
```

Here are the most downregulated lincRNAs (lowest to highest FC):

```{r}
head(DE_lincRNAs_asc)
```

Here is a list of lincRNAs organised from highest to lowest FDR:

```{r}
head(lincRNA_FDRs)
```

**Mt-rRNA**

```{r, include=FALSE}
Mt_rRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "Mt_rRNA")

Mt_rRNA <- Mt_rRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 
Mt_rRNA <- Mt_rRNA %>% mutate(log_FDR = -log10(FDR))

Mt_rRNA$Significance <- factor(Mt_rRNA$Significance, levels = c("Upreg","Downreg", "NS"))


Mt_rRNA_plot <- ggplot(data = Mt_rRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("Mt-rRNA")
```

```{r}
Mt_rRNA_plot
```

```{r, include=FALSE}
DE_Mt_rRNAs <- Mt_rRNA %>% select(Annotation, logFC, log_FDR)
DE_Mt_rRNAs_desc <- DE_Mt_rRNAs %>% arrange(desc(logFC))
DE_Mt_rRNAs_asc <- DE_Mt_rRNAs %>% arrange((logFC))

Mt_rRNA_FDRs <- DE_Mt_rRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated Mt rRNAs (highest to lowest FC):

```{r}
head(DE_Mt_rRNAs_desc)
```

Here are the most downregulated Mt rRNAs (lowest to highest FC):

```{r}
head(DE_Mt_rRNAs_asc)
```

Here is a list of Mt rRNAs organised from highest to lowest FDR:

```{r}
head(Mt_rRNA_FDRs)
```

**Mt-tRNA dysregulation**

```{r, include=FALSE}
Mt_tRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "Mt_tRNA")

Mt_tRNA <- Mt_tRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 
Mt_tRNA <- Mt_tRNA %>% mutate(log_FDR = -log10(FDR))

Mt_tRNA$Significance <- factor(Mt_tRNA$Significance, levels = c("Upreg","Downreg", "NS"))


Mt_tRNA_plot <- ggplot(data = Mt_tRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("Mt-tRNA")
```

```{r}
Mt_tRNA_plot
```

```{r, include=FALSE}
DE_Mt_tRNAs <- Mt_tRNA %>% select(Annotation, logFC, log_FDR)
DE_Mt_tRNAs_desc <- DE_Mt_tRNAs %>% arrange(desc(logFC))
DE_Mt_tRNAs_asc <- DE_Mt_tRNAs %>% arrange((logFC))

Mt_tRNA_FDRs <- DE_Mt_tRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated Mt tRNAs (highest to lowest FC):

```{r}
head(DE_Mt_tRNAs_desc)
```

Here are the most downregulated Mt tRNAs (lowest to highest FC):

```{r}
head(DE_Mt_tRNAs_asc)
```

Here is a list of Mt tRNAs organised from highest to lowest FDR:

```{r}
head(Mt_tRNA_FDRs)
```

**piRNA dysregulation**

```{r, include=FALSE}
piRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "piRNA")

piRNA <- piRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 

piRNA <- piRNA %>% mutate(log_FDR = -log10(FDR))

piRNA$Significance <- factor(piRNA$Significance, levels = c("Upreg","Downreg", "NS"))


piRNA_plot <- ggplot(data = piRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("piRNA")
```

```{r}
piRNA_plot
```

```{r, include=FALSE}
DE_piRNAs <- piRNA %>% select(Annotation, logFC, log_FDR)
DE_piRNA_desc <- DE_piRNAs %>% arrange(desc(logFC))
DE_piRNA_asc <- DE_piRNAs %>% arrange((logFC))

piRNA_FDRs <- DE_piRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated piRNAs (highest to lowest FC):

```{r}
head(DE_piRNA_desc)
```

Here are the most downregulated piRNAs (lowest to highest FC):

```{r}
head(DE_piRNA_asc)
```

Here is a list of piRNAs organised from highest to lowest FDR:

```{r}
head(piRNA_FDRs)
```

**SnoRNA dysregulation**

```{r, include= FALSE}
SnoRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "SnoRNA")

SnoRNA <- SnoRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 

SnoRNA <- SnoRNA %>% mutate(log_FDR = -log10(FDR))

SnoRNA$Significance <- factor(SnoRNA$Significance, levels = c("Upreg","Downreg", "NS"))


SnoRNA_plot <- ggplot(data = SnoRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("SnoRNA")
```

```{r}
SnoRNA_plot
```

```{r, include=FALSE}
DE_SnoRNAs <- SnoRNA %>% select(Annotation, logFC, log_FDR)
DE_SnoRNA_desc <- DE_SnoRNAs %>% arrange(desc(logFC))
DE_SnoRNA_asc <- DE_SnoRNAs %>% arrange((logFC))

SnoRNA_FDRs <- DE_SnoRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated SnoRNAs (highest to lowest FC):

```{r}
head(DE_SnoRNA_desc)
```

Here are the most downregulated SnoRNAs (lowest to highest FC):

```{r}
head(DE_SnoRNA_asc)
```

Here is a list of SnoRNAs organised from highest to lowest FDR:

```{r}
head(SnoRNA_FDRs)
```

**SnRNA dysregulation**

```{r, include=FALSE}
SnRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "SnRNA")

SnRNA <- SnRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        ))  

SnRNA <- SnRNA %>% mutate(log_FDR = -log10(FDR))

SnRNA$Significance <- factor(SnRNA$Significance, levels = c("Upreg","Downreg", "NS"))


SnRNA_plot <- ggplot(data = SnRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("SnRNA")
```

```{r}
SnRNA_plot
```

```{r, include=FALSE}
DE_SnRNAs <- SnoRNA %>% select(Annotation, logFC, log_FDR)
DE_SnRNA_desc <- DE_SnRNAs %>% arrange(desc(logFC))
DE_SnRNA_asc <- DE_SnRNAs %>% arrange((logFC))

SnRNA_FDRs <- DE_SnRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated SnRNAs (highest to lowest FC):

```{r}
head(DE_SnRNA_desc)
```

Here are the most downregulated SnRNAs (lowest to highest FC):

```{r}
head(DE_SnRNA_asc)
```

Here is a list of SnRNAs organised from highest to lowest FDR:

```{r}
head(SnRNA_FDRs)
```

**YRNA dysregulation**

```{r, include=FALSE}
YRNA <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "YRNA")

YRNA <- YRNA %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        ))  

YRNA <- YRNA %>% mutate(log_FDR = -log10(FDR))

YRNA$Significance <- factor(YRNA$Significance, levels = c("Upreg","Downreg", "NS"))


YRNA_plot <- ggplot(data = YRNA, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("YRNA")
```

```{r}
YRNA_plot
```

```{r, include=FALSE}
DE_YRNAs <- YRNA %>% select(Annotation, logFC, log_FDR)
DE_YRNA_desc <- DE_YRNAs %>% arrange(desc(logFC))
DE_YRNA_asc <- DE_YRNAs %>% arrange((logFC))

YRNA_FDRs <- DE_YRNAs %>% arrange(desc(log_FDR))

```

Here are the most upregulated YRNAs (highest to lowest FC):

```{r}
head(DE_YRNA_desc)
```

Here are the most downregulated YRNAs (lowest to highest FC):

```{r}
head(DE_YRNA_asc)
```

Here is a list of YRNAs organised from highest to lowest FDR:

```{r}
head(YRNA_FDRs)
```

**Dysregulation of sncRNAs with NO annotation**

```{r, include=FALSE}
NO_annotation <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "NO_Annotation")

NO_annotation <- NO_annotation %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 

NO_annotation <- NO_annotation %>% mutate(log_FDR = -log10(FDR))

NO_annotation$Significance <- factor(NO_annotation$Significance, levels = c("Upreg","Downreg", "NS"))


NO_annotation_plot <- ggplot(data = NO_annotation, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("NO Annotation")
```

```{r}
NO_annotation_plot
```

**Other**

```{r, include = FALSE}
Other <- read.xlsx("./1_data/Sch_edgeR_glmRT.xlsx", sheet = "other")

Other <- Other %>%
  mutate(Significance = if_else(FDR < FDR_threshold & logFC < 0 & logFC < -0.5833, "Downreg",
                           if_else(FDR < FDR_threshold & logFC > 0 & logFC > 0.5833, "Upreg", "NS")
                          
        )) 

Other <- Other %>% mutate(log_FDR = -log10(FDR))

Other$Significance <- factor(Other$Significance, levels = c("Upreg","Downreg", "NS"))


other_plot <- ggplot(data = Other, aes(x =logFC, y= log_FDR)) + geom_point(size = 2.5) + theme_minimal () + geom_hline(yintercept = -log10(FDR_threshold), linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = -0.5833, linetype = "dashed", col = "#999999") +
  geom_vline(xintercept = 0.5833, linetype = "dashed", col = "#999999") +
  aes(x = logFC, y = log_FDR, col = Significance, fill = Significance) +
  scale_color_manual(values = palette) + scale_fill_manual(values = palette) +
  xlab("log2FoldChange") + ylab("-log10(FDR)") + theme(legend.title = element_blank(), 
  axis.text = element_text(size = 14), axis.title = element_text(size = 18), 
  legend.text = element_text(size = 14)) + ggtitle("Other ")

```

```{r}
other_plot
```
